---
title: "Number metaphor analysis"
author: "Bodo"
date: "23/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

Load packages:

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
```

Define target nouns:

```{r}
nouns <- c('number', 'amount', 'proportion',
           'percentage', 'quantity', 'standard',
           'volume', 'cost', 'profit', 'income',
           'grade', 'loss', 'tax', 'benefit',
           'rate', 'gain', 'salary', 'pay',
           'revenue', 'earning')
```

Define adjectives combined with them. Do this for vertical and size based separately:

```{r}
verts <- c('high', 'higher', 'highest',
           'low', 'lower', 'lowest')
sizes <- c('small', 'smaller', 'smallest',
           'big', 'bigger', 'biggest',
           'large', 'larger', 'largest',
           'tiny', 'tinier', 'tiniest',   # sounds weird but for parallelism
           'huge', 'huger', 'hugest') # sounds weird but for parallelism

# Put them together:

adjs <- c(verts, sizes)
```

Do the same for deverbal adjectives (these tend to have a dynamic component):
(some inspiration from https://www.aclweb.org/anthology/W16-6606.pdf[https://www.aclweb.org/anthology/W16-6606.pdf])

```{r}
verts <- c('rising', 'falling',
           'climbing',
           'ascending', 'descending',
           'plunging', 'plummeting',
           'soaring', 'dipping', 'dropping',
           'slumping',
           'surging', 'skyrocketing')
neutral <- c('growing', 'shrinking',
             'increasing', 'decreasing')
deverbs <- c(verts, neutral)
```

## Setup objects for loop

Create empty data frame to be filled:

```{r}
# All adjective-noun combinations:

adj_targets <- expand.grid(adjs, nouns)

# Make to tibble, rename columns, and get a combined column:

adj_targets <- as_tibble(adj_targets) %>% 
  rename(adj = Var1, noun = Var2) %>% 
  mutate(adj_noun = str_c(adj, ' ', noun))

# Append empty frequency column to be filled:

adj_targets$freq <- numeric(nrow(adj_targets))

# Check:

sample_n(adj_targets, 10)
```

Setup target file with deverbal adjective-noun combinations:

```{r}
# All adjective-noun combinations:

deverb_targets <- expand.grid(deverbs, nouns)

# Make to tibble, rename columns, and get a combined column:

deverb_targets <- as_tibble(deverb_targets) %>% 
  rename(adj = Var1, noun = Var2) %>% 
  mutate(adj_noun = str_c(adj, ' ', noun))

# Append empty frequency column to be filled:

deverb_targets$freq <- numeric(nrow(deverb_targets))

# Check:

sample_n(deverb_targets, 10)
```

Get the file locations into a vector:

```{r}
spk_files <- list.files('../../COCA_spoken')

# Check:

spk_files
```

Extract just the totals for each term to compute PMI:

```{r}
totals <- tibble(word = c(adjs, nouns, deverbs))

# Empty frequency column:

totals$freq <- numeric(nrow(totals))

# Check:

sample_n(totals, 4)
```

And get the overall word counts:

```{r}
corpus_size <- tibble(file = spk_files)

# Empty frequency column:

corpus_size$freq <- numeric(nrow(corpus_size))

# Check:

sample_n(corpus_size, 4)
```

## Loop through files

```{r}
for (i in seq_along(spk_files)) {
  # Tell the outside world where you're at:
  
  cat(str_c('Starting file ', i, ' of ', length(spk_files), '\n'))
  
  # Get text file into character vector object:
  
  this_file <- suppressWarnings(readLines(str_c('../../COCA_spoken/', spk_files[i])))
  
  # # Save overall word count (the number of separators plus 1):
  # 
  # corpus_size[i, ]$freq <- sum(lengths(gregexpr('\\W+', this_file)) + 1)
  # 
  # # Save totals:
  
  for (j in 1:nrow(totals)) {
    this_total <- totals[j, ]$word
    this_sum <- sum(str_count(this_file, this_total))
    totals[j, ]$freq <- totals[j, ]$freq + this_sum
  }
  
  # # Get the frequencies of adjective-noun combinations:
  # 
  # for (k in 1:nrow(adj_targets)) {
  #   this_adj_noun <- adj_targets[k, ]$adj_noun
  #   this_sum <- sum(str_count(this_file, this_adj_noun))
  #   adj_targets[k, ]$freq <- adj_targets[k, ]$freq + this_sum
  # }
  # 
  # # Get the frequencies of deverbal noun phrase combinations:
  # 
  # for (k in 1:nrow(deverb_targets)) {
  #   this_adj_noun <- deverb_targets[k, ]$adj_noun
  #   this_sum <- sum(str_count(this_file, this_adj_noun))
  #   deverb_targets[k, ]$freq <- deverb_targets[k, ]$freq + this_sum
  # }
  # 
  
}
```

Write this file to folder:

```{r}
write_csv(totals, '../data/totals.csv')
write_csv(corpus_size, '../data/corpus_size.csv')
write_csv(adj_targets, '../data/adj_noun_spk_generic.csv')
write_csv(deverb_targets, '../data/deverb_noun_spk_generic.csv')
```



