---
title: "Play with metaphors"
author: "Bodo"
date: "23/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages and data:

Load packages:

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(ggcorrplot)
```

Load data:

```{r, warning = FALSE, message = FALSE}
totals <- read_csv('../data/totals.csv')
corpus_size <- read_csv('../data/corpus_size.csv')
adj_noun <- read_csv('../data/adj_noun_spk_generic.csv')
deverbs <- read_csv('../data/deverb_noun_spk_generic.csv')
```

Show each:

```{r}
# Totals (adjective and noun overall frequency):

sample_n(totals, 4)

# Size of individual files:

sample_n(corpus_size, 4)

# Adjective-noun combinations:

sample_n(adj_noun, 4)

# Adjective-noun combinations with deverbals:

sample_n(deverbs, 4)
```

These will have to be merged. We will use "adj_noun" as the main tibble to be analyzed.

## Merge and process data:

Merge totals into there:

```{r}
# Adjective-noun:

adj_noun$noun_N <- totals[match(adj_noun$noun, totals$word), ]$freq
adj_noun$adj_N <- totals[match(adj_noun$adj, totals$word), ]$freq

# Deverbal adjective-noun:

deverbs$noun_N <- totals[match(deverbs$noun, totals$word), ]$freq
deverbs$adj_N <- totals[match(deverbs$adj, totals$word), ]$freq
```

Mark whether an adjective is vertical or size-based:

```{r}
verts <- c('high', 'higher', 'highest',
           'low', 'lower', 'lowest')

adj_noun <- mutate(adj_noun,
                   type = ifelse(adj %in% verts, 'vertical', 'size'))
```

Show:

```{r}
sample_n(adj_noun, 4)
```

For the deverbal ones that's vertical vs. neutral:

```{r}
neutral <- c('growing', 'shrinking',
             'increasing', 'decreasing')

deverbs <- mutate(deverbs,
                  type = ifelse(adj %in% neutral, 'neutral', 'vertical'))
```

Show:

```{r}
sample_n(deverbs, 4)
```

## Compute PMI

Compute total frequency across the entire corpus. Will use this to normalize the frequencies:

```{r}
N <- sum(totals$freq)
```

Compute pointwise mutual information, which is:

log(p(x, y) / (p(x) * p(y)))

(from https://en.wikipedia.org/wiki/Pointwise_mutual_information[https://en.wikipedia.org/wiki/Pointwise_mutual_information])

First, we need the normalized frequencies:

```{r}
# Adjective-noun:

adj_noun <- adj_noun %>% 
  mutate(freq = ifelse(freq == 0, 0.1, freq)) %>% # not sure if this is legitimate
  mutate(p_x_y = freq / N,
         p_x = noun_N / N,
         p_y = adj_N / N,
         p_xy = p_x * p_y,
         pmi = p_x_y / p_xy,
         pmi = log(pmi)) %>% 
  select(-(p_x_y:p_xy))

# Deverbal adjective-noun:

deverbs <- deverbs %>% 
  mutate(freq = ifelse(freq == 0, 0.1, freq)) %>% # not sure if this is legitimate
  mutate(p_x_y = freq / N,
         p_x = noun_N / N,
         p_y = adj_N / N,
         p_xy = p_x * p_y,
         pmi = p_x_y / p_xy,
         pmi = log(pmi)) %>% 
  select(-(p_x_y:p_xy))
```

Check:

```{r}
sample_n(adj_noun, 10)
sample_n(deverbs, 10)
```

There are lots of zeros. This happens when the numerator is 0 and the log-transform is performed. Need to think about how to fix this.

Check the distribution of PMI:

```{r, fig.width = 4.5, fig.height = 3}
adj_noun %>% ggplot(aes(x = pmi)) +
  geom_density(col = 'black', fill = 'steelblue', alpha = 0.5) +
  theme_classic()
```

Same for deverbal:

```{r, fig.width = 4.5, fig.height = 3}
deverbs %>% ggplot(aes(x = pmi)) +
  geom_density(col = 'black', fill = 'steelblue', alpha = 0.5) +
  theme_classic()
```

## Average PMI for different nouns

Compute the average vertical / size PMI for different nouns:

```{r}
type_avgs <- adj_noun %>%
  filter(!is.infinite(pmi)) %>% 
  group_by(noun, type) %>% 
  summarize(pmi = mean(pmi, na.rm = TRUE))
```

Plot this:

```{r, fig.width = 12, fig.height = 6}
adj_p <- type_avgs %>% ggplot(aes(x = noun,
                         fill = type,
                         y = pmi)) +
  scale_fill_manual(values = c('steelblue', 'goldenrod2')) +
  geom_col(position = position_dodge(0.6),
           width = 0.6,
           col = 'black') +
  xlab(NULL) +
  theme_classic() +
  theme(legend.position = 'top',
        axis.text.x = element_text(angle = 45, hjust = 1, size = 14))

# Show and save:

adj_p
ggsave(plot = adj_p, filename = '../figures/adjectives.pdf',
       width = 16, height = 6)
```

Compute the average vertical / size PMI for different nouns for the deverbals:

```{r}
deverb_type_avgs <- deverbs %>%
  filter(!is.infinite(pmi)) %>% 
  group_by(noun, type) %>% 
  summarize(pmi = mean(pmi, na.rm = TRUE))
```

Plot this:

```{r, fig.width = 12, fig.height = 6}
deverb_p <- deverb_type_avgs %>% ggplot(aes(x = noun,
                         fill = type,
                         y = pmi)) +
  scale_fill_manual(values = c('steelblue', 'goldenrod2')) +
  geom_col(position = position_dodge(0.6),
           width = 0.6,
           col = 'black') +
  xlab(NULL) +
  theme_classic() +
  theme(legend.position = 'top',
        axis.text.x = element_text(angle = 45, hjust = 1, size = 14))

# Show and save:

deverb_p
ggsave(plot = deverb_p, filename = '../figures/deverb_adjectives.pdf',
       width = 16, height = 6)
```

## Create a relative PMI score

Create a relative PMI score that looks at the relative degree to which something is more vertical or more size-based (or conversely, more "neutral").

```{r}
rel_pmi_avgs <- type_avgs %>%
  pivot_wider(values_from = pmi, names_from = type) %>% 
  mutate(vertical_bias = vertical - size)
```

Make a plot of this:

```{r, fig.width = 12, fig.height = 6}
rel_pmi_p <- rel_pmi_avgs %>% 
  ggplot(aes(x = reorder(noun, vertical_bias), y = vertical_bias)) +
  geom_col(fill = 'salmon3') +
  xlab(NULL) +
  theme_classic() +
  theme(legend.position = 'top',
        axis.text.x = element_text(angle = 45, hjust = 1, size = 14))

# Show and save:

rel_pmi_p
ggsave(plot = rel_pmi_p, filename = '../figures/relative_pmi.pdf',
       width = 12, height = 6)
```

Same for the deverbal ones:

```{r}
deverb_rel_pmi_avgs <- deverb_type_avgs %>%
  pivot_wider(values_from = pmi, names_from = type) %>% 
  mutate(vertical_bias = vertical - neutral)
```

Make a plot of this:

```{r, fig.width = 12, fig.height = 6}
deverb_rel_pmi_p <- deverb_rel_pmi_avgs %>% 
  ggplot(aes(x = reorder(noun, vertical_bias), y = vertical_bias)) +
  geom_col(fill = 'salmon3') +
  xlab(NULL) +
  theme_classic() +
  theme(legend.position = 'top',
        axis.text.x = element_text(angle = 45, hjust = 1, size = 14))

# Show and save:

deverb_rel_pmi_p
ggsave(plot = deverb_rel_pmi_p, filename = '../figures/deverb_relative_pmi.pdf',
       width = 12, height = 6)
```

See whether the two are correlated:

```{r}
deverb_rel_pmi_avgs <- rename(deverb_rel_pmi_avgs,
                              deverb_bias = vertical_bias)

both_avgs <- rel_pmi_avgs %>% select(-size, -vertical) %>% 
  right_join(deverb_rel_pmi_avgs) %>% 
  select(-neutral, -vertical)
```

Make a plot of this:

```{r, fig.width = 8, fig.height = 6}
cor_p <- both_avgs %>% 
  ggplot(aes(x = deverb_bias, y = vertical_bias, label = noun)) +
  geom_smooth(method = 'lm', col = 'salmon3', fill = 'salmon4') +
  geom_text() +
  theme_classic()

# Show and save:

cor_p
ggsave(plot = cor_p, filename = '../figures/correlation.pdf',
       width = 6, height = 4)
```

## Check the correlation for individual lexical items within adjective set

If something prefers high, does it also prefer low; if something prefers, tiny, does it also prefer the other size terms etc.?

```{r}
# Lemmatize:

adj_noun <- mutate(adj_noun,
                   item = ifelse(adj %in% c('higher', 'highest'), 'high', adj),
                   item = ifelse(item %in% c('lower', 'lowest'), 'low', item),
                   item = ifelse(item %in% c('bigger', 'biggest'), 'big', item),
                   item = ifelse(item %in% c('smaller', 'smallest'), 'small', item),
                   item = ifelse(item %in% c('tinier', 'tiniest'), 'tiny', item),
                   item = ifelse(item %in% c('huger', 'hugest'), 'huge', item),
                   item = ifelse(item %in% c('larger', 'largest'), 'huge', item))
```

Get averages per item:

```{r}
item_avgs <- adj_noun %>% group_by(noun, item) %>% 
  summarize(pmi = mean(pmi)) %>% ungroup()

item_avgs
```

Make this into wide format for easier cross-correlations:

```{r}
item_avgs <- item_avgs %>% pivot_wider(values_from = pmi, names_from = item)

item_avgs
```

Cross-correlate and order them by vertical versus horizontal:

```{r}
cor_matrix <- cor(select(item_avgs, big:tiny))

cor_matrix <- cor_matrix[, c('high', 'low', 'big', 'small', 'huge', 'tiny', 'large')]
cor_matrix <- cor_matrix[c('high', 'low', 'big', 'small', 'huge', 'tiny', 'large'), ]
```

Show:

```{r}
round(cor_matrix, 2)
```

Make a plot of this:

```{r, fig.width = 8, fig.height = 6}
cor_p <- ggcorrplot(cor_matrix, type = "lower", hc.order = TRUE, 
           method = 'circle',
           outline.col = "white",
           ggtheme = ggplot2::theme_classic,
           colors = c("#6D9EC1", "white", "#E46726"))

cor_p
ggsave(plot = cor_p, filename = '../figures/cor_p.pdf')
```

Append rising and falling in there for now:

```{r}
big_df <- bind_rows(adj_noun, deverbs) %>%
  mutate(item = ifelse(adj %in% c('higher', 'highest'), 'high', adj),
         item = ifelse(item %in% c('lower', 'lowest'), 'low', item),
         item = ifelse(item %in% c('bigger', 'biggest'), 'big', item),
         item = ifelse(item %in% c('smaller', 'smallest'), 'small', item),
         item = ifelse(item %in% c('tinier', 'tiniest'), 'tiny', item),
         item = ifelse(item %in% c('huger', 'hugest'), 'huge', item),
         item = ifelse(item %in% c('larger', 'largest'), 'huge', item))
```

Get averages per item:

```{r}
item_avgs <- big_df %>% group_by(noun, item) %>% 
  summarize(pmi = mean(pmi)) %>% ungroup()

item_avgs
```

Make this into wide format for easier cross-correlations:

```{r}
item_avgs <- item_avgs %>% pivot_wider(values_from = pmi, names_from = item)

item_avgs
```

Cross-correlate and plot:

```{r, fig.width = 14, fig.height = 14}
cor_matrix <- cor(select(item_avgs, ascending:tiny))

cor_p <- ggcorrplot(cor_matrix, type = "lower", hc.order = TRUE, 
           method = 'circle',
           outline.col = "white",
           ggtheme = ggplot2::theme_classic,
           colors = c("#6D9EC1", "white", "#E46726"))

cor_p
ggsave(plot = cor_p, filename = '../figures/big_cor_p.pdf')
```

Show in markdown:

```{r}
round(cor_matrix, 2)
```


